The Ranked Nearest Neighbour Interchange space

### 2019-07-01 paper.tex
edit beginning of section geometry

This paper is an advance towards the general goal of understanding the geometry of the $\rnni$ space.
%TODO
In this section, we...

@@

This algorithm gives an upper bound for the diameter of $\rnni$, which improves the one given in \autocite{Gavryushkin2017}.
Later we will see that this bound is tight.

@@

In this section we restrict our attention to the set of \emph{caterpillar trees} where each internal node has at least one child that is a leaf.
This type of tree is of particular interest in $\rnni$ because shortest paths between such trees in $\rnni$ differ from those in classical $\nni$ space.
This in combination with the importance of caterpillar trees in the proof of $\np$-hardness of $\nni$ distances in \autocite{jiang2000} suggests that understanding the geometry of paths between caterpillar trees helps investigating the complexity of computing $\rnni$ distances.
In particular, we are going to prove that one can always find a shortest path between two caterpillar trees that contains caterpillar trees only, meaning that the set of caterpillar trees is convex in $\rnni$.
We discovered this by computing caterpillar paths using an algorithm to compute shortest paths within the set of caterpillar trees and comparing the results with shortest path in $\rnni$.
This algorithm will be explained in detail later in this section, after demonstrating why finding shortest paths between caterpillar trees builds the basis for understanding the difference between $\nni$ and $\rnni$.
Further details on our computational results can be found in section~\ref{computation}.

@@

\subsection{Diameter}

@@

\subsection{Split Theorem}

@@

\section{Computational results}
\label{computation}

>>>>

This paper is an advance towards the general goal of understanding the geometry of the $\rnni$ space.
In this section we are investigating the geometry by analysing the structure of shortest paths in $\rnni$.
We will start by presenting an algorithm for computing paths in $\rnni$ (Algorithm~\ref{alg:find_path}) that turns out to be an essential tool for our studies.
Using this algorithm, we are able to prove that there is always a shortest path between two caterpillar trees that only consists of caterpillar trees (Theorem~\ref{thm:caterpillar_convex}) in Section~\ref{section:caterpillar_convex}.
In other words, the set of caterpillar trees is convex in $\rnni$.
This distinguishes $\rnni$ from $\nni$ where this does not hold.
One further result, presented in Section~\ref{section:diameter}, directly following from the proof of Theorem~\ref{thm:caterpillar_convex}, is the diameter of $\rnni$.
Finally, we disprove the so-called Split Theorem, conjectured in \autocite{Gavryushkin2017}, and formulate an alternative version of it, the Cluster Theorem (Section~\ref{section:split_theorem}).

@@

This algorithm gives an upper bound for the diameter of $\rnni$, which improves the one given in \autocite{Gavryushkin2017}.
Later we will see that this bound is tight.
However, we don't know how good Algorithm~\ref{alg:find_path} distances in $\rnni$.
More details can be found in Section~\ref{section:computation}

@@

In this section we restrict our attention to the set of \emph{caterpillar trees} where each internal node has at least one child that is a leaf.
This type of tree is of particular interest in $\rnni$ because shortest paths between such trees in $\rnni$ differ from those in classical $\nni$ space.
This in combination with the importance of caterpillar trees in the proof of $\np$-hardness of $\nni$ distances in \autocite{jiang2000} suggests that understanding the geometry of paths between caterpillar trees helps investigating the complexity of computing $\rnni$ distances.
In particular, we are going to prove that one can always find a shortest path between two caterpillar trees that contains caterpillar trees only, meaning that the set of caterpillar trees is convex in $\rnni$.
We discovered this by computing caterpillar paths using an algorithm to compute shortest paths within the set of caterpillar trees and comparing the results with shortest path in $\rnni$.
This algorithm will be explained in detail later in this section, after demonstrating why finding shortest paths between caterpillar trees builds the basis for understanding the difference between $\nni$ and $\rnni$.
Further details on our computational results can be found in Section~\ref{section:computation}.

@@

\subsection{Diameter}
\label{section:diameter}

@@

\subsection{Split Theorem}
\label{section:split_theorem}

@@

\section{Computational results}
\label{section:computation}
